\section{Mixed Flagship Benchmark (C10) and Overall Results}
\label{sec:mixed_flagship}

\paragraph{The flagship workload.}
Benchmark C10 is the primary end-to-end evaluation workload, designed to exercise all layers of SheafDB simultaneously in a realistic multi-stage pipeline. It models a biomedical knowledge discovery scenario: given a set of seed entities, retrieve their interactors (neighborhood), filter by temporal validity (temporal), trace the provenance chain of each filtered fact (provenance), and verify global consistency across overlapping annotation sources (consistency).

\paragraph{Multi-stage pipeline.}
The pipeline proceeds in four stages, each corresponding to one of the targeted workload families:
\begin{enumerate}
  \item \textbf{Neighborhood (C1/C2)} --- retrieve all tuples mentioning any seed entity, with predicate filtering.
  \item \textbf{Temporal filter (C5)} --- restrict to tuples whose validity window intersects a query interval.
  \item \textbf{Provenance trace (C7)} --- walk the derivation chain for each surviving tuple to depth $d$.
  \item \textbf{Consistency check (C8)} --- verify that the provenance DAG satisfies consistency conditions across all overlapping annotation contexts.
\end{enumerate}

\paragraph{KG approach.}
A conventional RDF store must issue four separate SPARQL queries, each performing its own graph traversal. Intermediate results---sets of triple IDs---are materialized between stages and passed as SPARQL VALUES clauses or temporary tables.

\paragraph{SheafDB approach.}
SheafDB executes the full pipeline as a single optimized sheaf walk. The sheaf structure preserves entity locality (stage\,1), temporal indices are consulted during the walk without materializing intermediate results (stage\,2), provenance edges are followed via sheaf morphisms that share the walk state (stage\,3), and consistency is verified incrementally as each new context is visited (stage\,4). Shared state across stages avoids redundant computation.

\paragraph{Current status.}
The individual pipeline stages correspond to the per-workload benchmarks described in Sections~\ref{sec:event_reconstruction} through~\ref{sec:consistency_global}. Per-stage LOOKUP and GLOBAL query breakdowns appear in Section~\ref{sec:evaluation}. An integrated end-to-end pipeline that chains all four stages as a single benchmark is the subject of ongoing work; the present evaluation focuses on the per-stage characterisation that exposes the fundamental LOOKUP-vs-GLOBAL tradeoff.
\endinput
