Semantic knowledge is typically represented as collections of subject--predicate--object
triples. This representation is elegant for binary relationships, but it forces every
higher-arity fact and every context-scoped assertion to be decomposed through
reification, which multiplies the number of stored records and turns simple lookups into
multi-way joins. We investigate whether a database whose primary organising principle is
the context poset, rather than the triple, can offer a more direct representation for
facts that are inherently contextual or high-arity.

We describe the Semantic Fact Database (SFDB), a prototype system that stores each fact
as a section over a finite poset of contexts and evaluates queries through
stalk-indexed retrieval rather than triple reconstruction. The mathematical model is
drawn from the language of presheaves on finite posets equipped with the Alexandrov
topology. In this setting the sheaf condition holds vacuously, so we do not claim the
sheaf axioms as our technical contribution. What the model gives us is a disciplined
account of restriction, locality, and gluing that maps directly onto the operational
components of the database: the context index, the stalk index, and the global-section
constructor. In parallel we implement a conventional reified triple-store baseline over
the same core data model and prove semantic equivalence between the two engines through
a canonical intermediate representation, which every benchmark run verifies query by
query.

The evaluation compares both engines on four query classes
(LOOKUP, GLOBAL, bounded TEMPORAL, and unbounded TEMPORAL) plus insert throughput, at
scales of 100, 1\,000, and 10\,000 facts, with every query at every scale checked for
cross-engine result equivalence as part of the same run. Insert throughput is close to
parity: SFDB is $1.12\times$ slower than KG at $100$ facts, where absolute latencies
are small enough that the difference is noise-dominated, and $1.31$--$1.37\times$
faster at $1\,000$ and $10\,000$ facts. SFDB is faster than the KG baseline on every
query class measured: entity-anchored LOOKUP queries are faster by a factor of
$22$--$41$ because the stalk index returns complete facts in a single hash lookup
instead of reconstructing them from reified triples through joins; unrestricted
full-scan (GLOBAL) queries are faster by a factor of $39$--$95$ once the query path is
changed to walk the same flat stalk index rather than enumerate topology open sets;
bounded temporal-interval queries are faster by a factor of $1.4$--$1.7$ via a
year-bucketed index; and unbounded temporal-interval queries (open-ended ranges with
no upper bound, which the bucketed index alone cannot narrow) are faster by a factor
of $1.1$--$1.4$ once resolved against a second, flat, binary-searchable temporal
index built alongside the bucketed one specifically for this case --- a limitation we
originally reported as an open architectural gap and have since closed, described in
Section~\ref{sec:discussion:temporal-unbounded}. This is not a uniformly favourable
result, however: SFDB's in-memory indexing structure uses $1.2$--$3.4\times$ more
resident memory per fact than the KG baseline's more compact reified-triple storage,
and we report this tradeoff plainly. We also report, as a direct consequence of
tightening the cross-engine verification harness during this evaluation, that an
earlier version of the sheaf engine silently answered entity-anchored LOOKUP queries
with neighbourhood-style (any-mention) semantics rather than strict subject match; the
stricter verification caught the discrepancy before publication, which we take as
evidence for the value of the equivalence-checking methodology this paper describes,
not merely for the numbers it produced.

The contribution of this paper is therefore not a claim of universal performance
superiority nor a novel piece of category theory. It is (i)~a workable design for a
context-indexed semantic fact database backed by an implementation and full test suite,
(ii)~a mechanised proof of semantic equivalence between the sheaf-inspired engine and a
conventional triple store, enforced by an automated verification harness that we show
catches real cross-engine semantic bugs, and (iii)~a careful benchmark that maps the
design's latency advantages against its memory cost across a LOOKUP/GLOBAL/TEMPORAL
taxonomy, including a case study in closing a genuine architectural gap (unbounded
temporal ranges) once it was identified. All results are reproducible from the
released artifact with a single command.
