\section{Complexity Analysis}
\label{sec:complexity}

All analyses assume the standard RAM model with uniform cost for basic
operations (hash lookup, pointer dereference, arithmetic).

\paragraph{Notation:}
$N$ = number of facts,
$k$ = fact arity,
$|\mathcal{C}|$ = number of contexts,
$d$ = context depth in the poset,
$T$ = number of KG triples,
$R$ = number of restriction maps,
$n$ = result size.

\subsection{Fact Construction and Modification}
\label{sec:complexity:modification}

Constructing a fact requires building the object tuple:
time $O(k)$, space $O(k + m)$ where $m$ is metadata size.

\begin{table}[ht]
\centering
\caption{Insertion and deletion complexity.}
\label{tab:complexity:modification}
\begin{tabular}{@{}lcc@{}}
\toprule
Operation & Knowledge Graph & Sheaf Database \\
\midrule
Insert & $O(k \log T)$ & $O(d)$ amortized, $O(|\mathcal{C}|)$ worst \\
Delete & $O(k \log T)$ & $O(|\mathcal{C}|)$ worst \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{KG insertion} decomposes a fact into $2+k$ reified triples and
inserts each into three SPO/POS/OSP indices, costing $O(k \log T)$.

\paragraph{Sheaf insertion} stores the fact in its context section ($O(1)$ hash)
and registers it with open sets. With context chain traversal this is $O(d)$
amortized, $O(|\mathcal{C}|)$ worst case.

\subsection{Query Operations}
\label{sec:complexity:query}

\begin{table}[ht]
\centering
\caption{Query complexity comparison. $T$ = number of KG triples, $y$ = year
buckets spanned by a temporal range, $n_y$ = mean candidates per bucket.
See the paragraphs below for the reasoning behind each entry.}
\label{tab:complexity:query}
\begin{tabular}{@{}lcc@{}}
\toprule
Operation & Knowledge Graph & Sheaf Database \\
\midrule
Entity Lookup & $O(\log T + n)$ & $O(1 + n)$ \\
Full Scan (GLOBAL) & $O(T)$ & $O(N)$ \\
Context Lookup & $O(\log T + n)$ & $O(\log |\mathcal{C}| + n)$ \\
Restriction & N/A (triple-oriented) & $O(k)$ \\
Global Sections (Gluing) & N/A & $O(|\mathcal{C}| \cdot N^2)$ worst \\
Neighbourhood Search & $O(\log T + n)$ & $O(d + n)$ \\
Temporal Range (bounded) & $O(T)$ & $O(y \cdot n_y)$ \\
Temporal Range (unbounded) & $O(T)$ & $O(N)$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Entity Lookup.} KG locates every triple mentioning the entity through
its POS index and reassembles each matching fact from its reified triples.
SFDB resolves the entity to a candidate fact-id set via the neighbourhood
index in expected $O(1)$ per candidate, then returns a stalk-index lookup per
candidate fact --- also $O(1)$ each, since every stalk holds exactly one
section. Both engines are implemented as genuine hash-based lookups; the
gap observed empirically (Section~\ref{sec:eval:query}) is a constant-factor
one, set by SQLite round-trips and reified-triple reconstruction on the KG
side rather than by asymptotic complexity.

\paragraph{Full Scan (GLOBAL).} An unrestricted query that must return every
fact. KG scans its triple table once ($O(T)$ where $T = O(N \cdot k)$ is the
number of reified triples) and reconstructs each fact from its triples.
SFDB iterates the flat stalk index directly ($O(N)$), since every fact is
registered in exactly one stalk regardless of how many topology open sets it
also belongs to. An earlier iteration of this design answered GLOBAL queries
by enumerating every open set a fact could appear in and de-duplicating,
which cost $O(N \cdot \bar{o})$ where $\bar{o} \approx 8$--$10$ is the mean
number of open sets per fact --- correct but strictly dominated by the flat
index that was already being maintained for entity lookups. We describe this
correction in Section~\ref{sec:discussion:interpretation}.

\paragraph{Context Lookup.} KG requires an index scan; SFDB performs a direct
open set lookup with context-to-open-set index.

\paragraph{Restriction.} Given a fact $f$ with arity $k$ and contexts $c \leq d$:
filter object slots meaningful in $c$ ($O(k)$) and specialize values ($O(k)$).
Total $O(k)$, independent of $N$.

\paragraph{Global Sections (Gluing).} This is a distinct operation from the
GLOBAL query class benchmarked in Section~\ref{sec:evaluation}: it computes
and verifies pairwise compatibility of sections across overlapping contexts
(the $\textsc{Glue}()$ operation of Section~\ref{sec:problem:formal}), rather
than simply returning every stored fact. The algorithm processes the context
poset bottom-up:
\begin{enumerate}
    \item Leaf collection: scan all facts at minimal contexts --- $O(N)$.
    \item Bottom-up gluing: for each non-leaf context $c$, check all
          sub-context pairs. Worst case (fully-connected lattice with
          incompatible sections): $O(|\mathcal{C}| \cdot N^2)$.
    \item Expected case (tree-structured contexts, most sections compatible):
          $O(N \log |\mathcal{C}|)$.
\end{enumerate}
This operation is not exercised by the LOOKUP, GLOBAL, or TEMPORAL query
classes reported in Section~\ref{sec:evaluation}, so its cost is not
reflected in those measurements; it remains a genuine scalability concern
for workloads that call $\textsc{Glue}()$ directly, and is unverified beyond
the unit-test scale reported in Section~\ref{sec:correctness}.

\paragraph{Neighbourhood Search.} Given a point $x$ with minimal neighbourhood
$\mathcal{B}(x) = \downarrow \alpha(x)$, compute $\mathcal{B}(x)$ in $O(1)$,
expand outward by $d$ steps, and collect sections in $O(n)$ cumulative.
Total: $O(d + n)$.

\paragraph{Temporal Search.} A range query bounded on both ends narrows to the
$y$ year buckets it spans via the temporal index, each holding $n_y$
candidate facts on average, then filters each candidate against the exact
interval in $O(1)$: $O(y \cdot n_y)$ total, which is sub-linear in $N$ for a
narrow range. An open-ended range (no upper bound) cannot be narrowed this
way and falls back to consulting every year bucket the index has ever
created, degrading to $O(N)$ --- empirically a $\sim 40\times$ slowdown
relative to the bounded case at $N=10^4$
(Section~\ref{sec:discussion:temporal-unbounded}). KG has no temporal index;
it always scans the triple table and filters reconstructed facts against
the query range in Python, $O(T)$ regardless of whether the range is
bounded.

\subsection{Space Complexity}
\label{sec:complexity:space}

\begin{table}[ht]
\centering
\caption{Space complexity comparison.}
\label{tab:complexity:space}
\begin{tabular}{@{}lcc@{}}
\toprule
Component & KG & Sheaf Database \\
\midrule
Per fact & $O(k \cdot 3)$ triples $+$ indices & $O(1)$ section $+$ $O(d)$ restriction entries \\
Context poset & $O(1)$ (not stored) & $O(|\mathcal{C}|^2)$ for topology \\
Restriction maps & N/A & $O(R) \leq O(|\mathcal{C}| \cdot N)$ \\
Indices & $O(T)$ via SPO/POS/OSP & $O(N + |\mathcal{C}|)$ via context hash \\
\bottomrule
\end{tabular}
\end{table}

KG stores each reified fact as $2+k$ triples, each indexed in up to three
lookup tables --- a $3\times$ index amplification. SFDB stores each fact once
in its context section plus $O(d)$ restriction entries for the ancestors;
context indexing adds $O(|\mathcal{C}|^2)$ for the topology and
$O(N + |\mathcal{C}|)$ for the context hash.

\subsection{Empirical Hypotheses}
\label{sec:complexity:hypotheses}

These bounds are tested empirically by the benchmark framework
(Section~\ref{sec:evaluation}):

\begin{enumerate}
    \item SFDB insertion is typically $O(d)$ vs.\ KG's $O(k \log T)$ for contexts
          of depth $d$; empirically SFDB matches or slightly beats KG insert
          latency at every scale up to $10^4$ facts (up to $1.35\times$
          faster, narrowing to near-parity at $100$ facts).
    \item SFDB LOOKUP queries are faster than KG's reified-triple
          reconstruction at every scale measured ($21$--$32\times$ across
          $10^2$--$10^4$ facts), consistent with both being $O(1)$-per-candidate
          lookups with a KG-side constant-factor penalty from SQLite round trips.
    \item SFDB GLOBAL queries are faster than KG's triple-table scan at
          every scale measured ($44$--$95\times$ across $10^2$--$10^4$
          facts), once the query path uses the flat stalk index rather
          than open-set enumeration; this reverses what an earlier, buggy
          implementation measured, as discussed in
          Section~\ref{sec:discussion:interpretation}.
    \item SFDB's temporal index gives a modest speedup ($1.2$--$1.6\times$)
          on bounded ranges but degrades below KG's flat scan-and-filter on
          unbounded ranges, exactly as the $O(y \cdot n_y)$ vs.\ $O(N)$
          bounds above predict.
    \item The expected-case $O(N \log |\mathcal{C}|)$ gluing complexity for
          the $\textsc{Glue}()$ operation (distinct from a GLOBAL query) is
          conjectured for tree-structured context hierarchies and is
          checked only at unit-test scale; it is not benchmarked at
          $10^4$ facts in this paper.
\end{enumerate}
