\section{Conclusion}
\label{sec:conclusion}

We described the Semantic Fact Database, a prototype context-indexed
storage system for semantic facts. The system stores each fact atomically
as a section over a finite context poset and answers entity-anchored
queries through a stalk index that returns complete facts in a single hash
lookup. A companion reified triple-store baseline is provided over the
same canonical schema, and a mechanised proof of semantic equivalence
binds the two engines together: every insert and every query is checked,
online, for canonical result equality.

The evaluation, on four query classes and insert throughput at three
scales, yields a clear result and a clear tradeoff. SFDB is faster on
every measured dimension: inserts by $1.05$--$1.36\times$; entity-anchored
LOOKUP queries by $23$--$32\times$ because the stalk index removes the need
for triple reassembly; unrestricted GLOBAL queries by $44$--$95\times$ once
that same flat index is used instead of enumerating topology open sets;
bounded temporal-interval queries by $1.2$--$1.4\times$ via the
engine's year-bucketed temporal index; and unbounded temporal-interval
queries by $1.1$--$1.2\times$ via a second, flat,
binary-searchable temporal index we added specifically because the
bucketed index alone cannot narrow an open-ended range. The tradeoff that
funds these advantages is memory: SFDB uses $1.3$--$2.5\times$ more
resident memory per fact than the reified-triple baseline at the two
larger scales measured. We report this
plainly, and it is now the paper's only reported latency-independent
tradeoff, following the fix described above. We also report four bugs and
performance gaps found and fixed while preparing this evaluation, by three
different means: the sheaf engine's LOOKUP path silently used
neighbourhood (any-mention) rather than strict-subject match, caught by
tightening the cross-engine verification harness itself
(Section~\ref{sec:correctness}); the GLOBAL open-set enumeration
inefficiency that produced an earlier draft's headline negative result was
found by inspecting the code path, not by the harness, since both engines
already agreed on results before the fix and disagreed only on speed
(Section~\ref{sec:discussion:interpretation}); a linear-scan identifier
decoder in the KG baseline, also found by inspection, which had been
inflating the KG engine's own reconstruction cost; and the unbounded
temporal-range degradation, found by the same code-inspection route as the
GLOBAL fix and closed the same way, by adding an index the design already
had room for rather than by redesigning around the limitation
(Section~\ref{sec:discussion:temporal-unbounded}). We treat the fact that
our own evaluation needed four rounds of correction, by three different
mechanisms, as part of the paper's evidence for the value of that scrutiny
rather than as an embarrassment to suppress.

The formal apparatus in the paper is that of presheaves on a finite poset.
We do not claim a novel sheaf-theoretic theorem: on the Alexandrov topology
that fits our setting, the sheaf condition holds vacuously and could not
serve as a technical contribution. The value of the framework is
operational. It gives us a disciplined vocabulary for restriction, stalks,
and global sections that maps directly onto the components of the
implementation, and it gives us a canonical intermediate representation
that makes cross-engine equivalence a checkable property rather than an
aspiration.

We do not claim that the design supplants production RDF stores. Our KG
baseline is a controlled research implementation and the reported speedups
are best interpreted as a measurement of what disciplined context-indexed
storage buys over a strict reification-plus-joins design at comparable
engineering effort. Extending the evaluation to standard benchmarks and to
production RDF endpoints is future work. The artifact and reproducibility
manifest for every result in the paper are released with the paper.
