\section{Discussion}
\label{sec:discussion}

\subsection{What the Results Do and Do Not Show}
\label{sec:discussion:interpretation}

The measurements in Section~\ref{sec:evaluation} show SFDB faster than the
KG baseline on every query class measured: $22$--$41\times$ on LOOKUP,
$39$--$95\times$ on GLOBAL, $1.4$--$1.7\times$ on bounded TEMPORAL queries,
and $1.1$--$1.4\times$ on unbounded TEMPORAL queries, at close-to-parity
insert latency. This was not the paper's original finding, on two separate
occasions. An earlier iteration of this evaluation reported GLOBAL queries
as SFDB's weak point --- up to three orders of magnitude slower than the KG
baseline --- because the query path enumerated topology open sets rather
than a flat fact index. That earlier finding was correct as a measurement
and wrong as an architectural conclusion: the open-set enumeration was an
implementation choice, not a consequence of context-indexing, and a flat
stalk index the engine already builds for LOOKUP queries answers GLOBAL
queries directly once the query path is changed to use it (a change of
roughly fifteen lines). A second, later iteration reported unbounded
TEMPORAL queries as a genuine architecture-level limitation rather than a
fixable implementation gap; that conclusion was also wrong, for the same
underlying reason, and Section~\ref{sec:discussion:temporal-unbounded}
describes the fix. We describe both earlier findings and their fixes here
rather than silently updating the numbers, because the fact that the
original measurements were real but the conclusions drawn from them were
not is itself informative: an implementation-level gap in a code path can
look, from the outside, exactly like a fundamental tradeoff until someone
checks whether a cheaper path can be built alongside the existing one.

The tradeoff this paper does report, and that survives this scrutiny, is
memory: SFDB uses $1.2$--$3.4\times$ more resident memory per fact than the
KG baseline (Section~\ref{sec:eval:memory}). This is a structural
consequence of the design, not an implementation oversight --- every fact
is registered in roughly eight to ten open sets plus four auxiliary
indexes, and each registration holds a live reference into memory rather
than a compact encoded row. Reducing it would require either coarsening
the topology (fewer open sets per fact) or compressing the index
structures (Section~\ref{sec:future:compression}), both of which would
likely give back some of the latency advantage.

We do not claim, and the data do not support, that the sheaf-inspired
approach dominates a mature RDF store on every workload at every scale. Our
KG baseline is a research prototype implemented by the same team as the
sheaf engine and running against the same test harness, storage backend,
and canonicalisation pipeline. It is a controlled reference, not a
state-of-the-art. Any claim about how SFDB would fare against Jena TDB2 or
Virtuoso must wait for a comparable evaluation, which we identify as future
work.

\subsection{Fixing the Unbounded Temporal Range}
\label{sec:discussion:temporal-unbounded}

The bounded TEMPORAL results in Section~\ref{sec:evaluation} use a query
bounded on both ends (a one-year window), which is the case the
year-bucketed temporal index is built for. An open-ended query (a start
bound with no end, "everything from year $X$ onward") cannot be narrowed
to a bounded set of year buckets, so a classifier that only ever consults
that index has to fall back to every year it has ever seen. An earlier
version of this evaluation measured exactly that: at $10^4$ facts, SFDB's
TEMPORAL latency degraded by roughly $40\times$ relative to the bounded
case, to the point of being slower than the KG baseline rather than
faster, and we reported it as a second genuine, architecture-level
limitation, distinct from the GLOBAL enumeration bug described above ---
specifically, we wrote that it was "not a fifteen-line fix, because there
is no bounded candidate set to compute."

That claim does not survive scrutiny any better than the GLOBAL one did.
There is no bounded candidate set within the \emph{year-bucketed} index,
but nothing requires the fix to live in that index. We added a second,
flat structure to \texttt{sfdb.sheaf.indexes.TemporalIndex}: fact ids
sorted by temporal end timestamp, alongside a separate set for facts with
no end (still valid, and therefore always overlapping any open-ended
range). A query
with a start bound and no end bound binary-searches this sorted array for
the first entry ending after the bound and takes everything from there
onward, unioned with the always-valid set --- $O(\log M + n)$ in the
number of temporally-annotated facts $M$, where $n$ is the candidate set
size, rather than $O(N)$ for consulting every bucket. (The symmetric case,
an end bound with no start bound, uses a mirror structure sorted by start
timestamp.) This is exactly the design this paper's future-work section
had proposed as the fix for this gap
(Section~\ref{sec:future:global-index} described it before we implemented
it); building it turned out to take an afternoon, not a research
programme.

With the flat index in place, unbounded TEMPORAL queries are
$1.1$--$1.4\times$ faster on SFDB than on the KG baseline
(Section~\ref{sec:eval:query}), which no longer scans every reified
temporal triple with an advantage over SFDB's fallback, because SFDB no
longer has a $O(N)$ fallback to fall back to. The advantage is smaller
than on bounded TEMPORAL queries because both engines still do comparable
per-candidate interval-overlap filtering once the candidate set is
identified; narrowing that candidate set further (e.g.\ an interval tree
that also incorporates the end bound when one is present) is a smaller
optimisation than the one made here, and we did not chase it. We record
the history here, alongside the fix, in the same spirit as
Section~\ref{sec:discussion:interpretation}: the original measurement was
real, the "architectural" framing around it was not, and only checking
whether a cheap structural fix existed revealed the difference.

\subsection{When the Design Helps}
\label{sec:discussion:sheaf-wins}

The context-indexed design helps when the workload has any of the
following properties. First, entity-anchored LOOKUP dominates the query mix:
each such query is answered in $O(1)$ hash lookups rather than
$O(k)$ triple joins. Second, facts have arity above two or three: reification
overhead grows linearly with arity in the KG baseline, while the SFDB engine
stores each fact atomically. Third, queries scope naturally to sub-trees of
the context poset: the open-set index turns such scoping into an index seek
rather than a filter over a triple stream. Fourth, the application requires
verifiable data integrity across representations: the canonical layer and
cross-engine verification harness make it cheap to run the same fact
stream through both engines and catch inconsistencies. Fifth, and
contrary to our own earlier expectation, unrestricted full-table scans
are not actually a case where the KG baseline has an advantage once the
GLOBAL query path uses the same flat index as LOOKUP. Sixth, and also
contrary to an earlier expectation of ours, open-ended temporal ranges are
not a case where the KG baseline has an advantage either, once a second
flat index is built alongside the year-bucketed one
(Section~\ref{sec:discussion:temporal-unbounded}); the design constraint
that remains, across every class we measured, is memory, not access
pattern.

\subsection{When the KG Baseline Helps}
\label{sec:discussion:kg-wins}

The reified triple representation still wins in workloads that the SFDB
design was not built for, even though it no longer wins on latency for any
of the four classes reported in Section~\ref{sec:evaluation} --- including
the open-ended temporal range that an earlier version of this evaluation
identified as a concrete case where SFDB's indexing advantage degraded and
the KG baseline was no worse (Section~\ref{sec:discussion:temporal-unbounded}
describes the fix). Workloads that are essentially
binary---where facts genuinely are subject-predicate-object triples with
no context, no arity, and no provenance---get no benefit from the sheaf
structure and only pay the memory overhead of maintaining the extra
indexes documented in Section~\ref{sec:eval:memory}, for no corresponding
latency win. Any workload where memory is the binding constraint rather
than latency favours the KG baseline's more compact reified storage.
Finally, integration with the extensive existing RDF and SPARQL tooling
ecosystem is currently only possible on the KG side.

\subsection{Threats to Validity}
\label{sec:discussion:threats}

We identify four threats that a careful reader should weigh.

\begin{description}
    \item[Implementation bias.] Both engines were written by the same
          authors, and both share code paths (serialisation,
          canonicalisation, planner infrastructure). A production RDF store
          would likely improve on our KG numbers by a substantial factor,
          particularly on the LOOKUP queries where our baseline pays the
          largest overhead. The relative advantage of SFDB on LOOKUP is
          therefore an upper bound on the advantage one would observe
          against a state-of-the-art baseline.

    \item[Dataset bias.] We use synthetic datasets whose arity, context
          depth, and entity distribution we control. Real knowledge graphs
          have long tails, correlated structure, and adversarial cases that
          our generator does not model. Repeating the evaluation on
          LUBM~\cite{guo2005lubm} and DBpedia is the natural next step.

    \item[Query bias.] Our query set was designed with both engines in
          mind: one anchor-based class (LOOKUP), one no-anchor class
          (GLOBAL), and one range-based class measured under two parameter
          settings (bounded and unbounded TEMPORAL). A different query mix
          would shift the aggregate numbers; Section~\ref{sec:discussion:temporal-unbounded}
          shows that varying a single parameter of the TEMPORAL class
          (whether the range is bounded) used to be enough to flip which
          engine is faster, before we added the index that closes that gap.
          We do not claim the four classes measured here are exhaustive, or
          that no query shape exists under which the KG baseline's simpler
          scan-and-filter approach would win on latency --- only that none
          of the four we measured show one.

    \item[Scale.] All measurements are at $10^4$ facts or below. Standard
          RDF benchmarks start at $10^5$ triples. Our design choices are
          consistent with the complexity analysis in Section~\ref{sec:complexity},
          but empirically our claims are bounded at $10^4$.
\end{description}

We also record, in the interest of showing our work rather than only its
conclusion, that four bugs and performance gaps were found and fixed
during this evaluation, by three different mechanisms, only one of which
was the evaluation methodology catching itself automatically:

\begin{enumerate}
    \item The sheaf engine's LOOKUP path silently used neighbourhood
          (any-mention) match semantics instead of strict subject match.
          This one genuinely was caught by the cross-engine verification
          harness described in Section~\ref{sec:correctness}, after we
          tightened it from a positional, stringified comparison to an
          order-independent, type-preserving one: the two engines'
          canonical result sets stopped agreeing, which is exactly what
          the harness is for.
    \item The GLOBAL open-set enumeration inefficiency that produced an
          earlier draft's headline negative result (Section~\ref{sec:discussion:interpretation})
          was \emph{not} caught by the harness, because both engines
          already agreed on which facts a GLOBAL query returned before the
          fix --- they disagreed only on how long it took. It was found by
          inspecting the query path and noticing a cheaper index was
          already being maintained elsewhere for LOOKUP.
    \item A linear scan standing in for a dictionary lookup in the KG
          engine's identifier decoder was found the same way, by code
          inspection rather than by any automated check, and had been
          inflating the KG engine's own reconstruction cost on every
          query class that touches reified triples.
    \item The unbounded-TEMPORAL degradation described in
          Section~\ref{sec:discussion:temporal-unbounded} was found and
          fixed by the same mechanism as the GLOBAL case: code inspection
          revealed that a second flat index, sorted and binary-searchable,
          could sit alongside the existing year-bucketed one at a fraction
          of the engineering cost the paper had previously estimated for a
          fix. Like the GLOBAL fix, this was not caught by the harness
          either, since it changed only how long a correct answer took to
          compute.
\end{enumerate}

All four fixes are reflected in the numbers reported in
Section~\ref{sec:evaluation}. That only the first was actually caught by
an automated check is itself a threat to validity worth naming plainly:
the verification harness is real and it does its job, but it checks
result equivalence, not performance, and it cannot catch an inefficiency
that both engines agree is merely slow. A less careful reading of the
same benchmark output --- one that trusted the harness to have caught
everything worth catching, or that trusted an earlier draft's own
"architectural limitation" framing at face value --- would have reported
the pre-fix numbers as a stable finding, twice.

\subsection{Negative Results}
\label{sec:discussion:negative}

Two negative results deserve to be stated plainly, because they discipline
the paper's overall claim. An earlier version of this section reported
three: the third was the degradation of TEMPORAL performance under an
open-ended range, which inverted SFDB's bounded-range advantage into a
slowdown because the year-bucketed index had no way to narrow an unbounded
query. Section~\ref{sec:discussion:temporal-unbounded} describes the fix
that closed it (a second, flat, binary-searchable temporal index), and
Section~\ref{sec:eval:query} reports SFDB now $1.1$--$1.4\times$ faster on
exactly this query shape. We keep the history here rather than deleting
the paragraph, for the same reason the fix itself is described in detail
elsewhere in this section: a negative result that turns out to be a fixed
implementation gap rather than an architectural ceiling is worth recording
as such, not quietly dropped once resolved.

The first of the two negative results that remain is the memory cost
quantified in Section~\ref{sec:eval:memory}: SFDB uses $1.2$--$3.4\times$
more resident memory per fact than the KG baseline. This is the tradeoff
that funds every latency advantage reported in this paper, and it is
intrinsic to the design rather than an implementation artefact: redundant
per-open-set registration is what makes the stalk and temporal indexes
cheap to query.

The second is the vacuousness of the sheaf condition in our setting
(Section~\ref{sec:sheaf-remark}): the system is, strictly, a presheaf
database with a defensive consistency check, not a sheaf database in any
technical sense that would constitute a contribution on its own. This does
not diminish the practical value of the design: what makes the design work
is the context poset, the stalk index, the canonical intermediate
representation, and the disciplined restriction vocabulary that comes with
them. It does mean that any framing of the paper
as a sheaf-theoretic breakthrough would be dishonest, and we do not adopt
that framing.
