% Evaluation — verified benchmark results
\section{Evaluation}
\label{sec:evaluation}

\subsection{Experimental Setup}

All experiments were run on a single machine (Windows 11, Python 3.14)
with both engines using in-memory SQLite storage.  Each data point
is the average of 10 warm-cache runs after 2 warm-up iterations.
The synthetic dataset generator creates facts with controlled arity,
context depth, and entity distribution (seed 42).

\subsection{Insert Performance}

Table~\ref{tab:insert} shows insert throughput at three scales.
Both engines achieve comparable insert performance, with SheafDB
slightly faster due to its simpler storage model (no triple decomposition).

\begin{table}[htbp]
\centering
\begin{tabular}{lrrr}
\toprule
Facts & KG (ms) & SheafDB (ms) & Ratio \\
\midrule
100   & 5.8     & 5.1          & 0.88$\times$ \\
1,000 & 59.0    & 54.1         & 0.92$\times$ \\
10,000& 651.1   & 627.0        & 0.96$\times$ \\
\bottomrule
\end{tabular}
\caption{Insert latency (lower is better).  SheafDB avoids triple
decomposition overhead, achieving slightly faster inserts.}
\label{tab:insert}
\end{table}

\subsection{Query Performance}

Table~\ref{tab:query} reports query latency for LOOKUP and GLOBAL
queries at three scales.  SheafDB's LOOKUP is orders of magnitude
faster because it uses a hash-indexed stalk for O(1) fact retrieval,
while KG must reconstruct facts from reified triples via SQL joins.
GLOBAL queries (full scans) are faster on KG because the Sheaf
engine iterates over all open sets.

\begin{table}[htbp]
\centering
\begin{tabular}{lrrrr}
\toprule
Query & Facts & KG (ms) & SheafDB (ms) & Ratio \\
\midrule
LOOKUP & 100   & 0.008 & 0.003  & 0.32$\times$ \\
LOOKUP & 1,000 & 0.520 & 0.007  & 0.01$\times$ \\
LOOKUP & 10,000& 2.279 & 0.074  & 0.03$\times$ \\
\midrule
GLOBAL & 100   & 3.359 & 3.244  & 0.97$\times$ \\
GLOBAL & 1,000 & 3.718 & 91.081 & 24.50$\times$ \\
GLOBAL & 10,000& 7.403 & 10407  & 1405$\times$ \\
\bottomrule
\end{tabular}
\caption{Query latency (lower is better).  SheafDB excels at
entity LOOKUP via stalk index but struggles with full scans
due to open set iteration overhead.}
\label{tab:query}
\end{table}

\subsection{Cross-Engine Verification}

All benchmark runs passed cross-engine verification: both KG and
SheafDB produce identical canonical results for every query.
The verification framework compares result sets by fact ID,
subject, relation, objects, and context, confirming semantic
equivalence across representations.

\subsection{Summary of Findings}

\begin{enumerate}
    \item \textbf{Insert:} SheafDB $\approx$ KG.  Both engines scale
          linearly with dataset size at $\sim$60$\mu$s/fact.
    \item \textbf{LOOKUP:} SheafDB $\gg$ KG.  Stalk-indexed retrieval
          is O(1) vs.\ KG's O(k) triple reconstruction.
    \item \textbf{GLOBAL:} KG $\gg$ SheafDB.  Full scans expose
          SheafDB's open set iteration overhead.
    \item \textbf{Verification:} All cross-engine comparisons pass.
          The canonical model guarantees semantic equivalence.
\end{enumerate}

\subsection{Limitations}

The evaluation compares SheafDB against its own KG baseline, not
against production RDF stores (Jena, GraphDB, Virtuoso).  The Jena
adapter has partial rdflib support but is not yet integrated into
the benchmark pipeline.  Results at 10K facts are indicative but
do not demonstrate production-scale behavior.
