\section{Future Work}
\label{sec:future}

Several concrete extensions would improve either the design or the
credibility of the evaluation. We list them in the order we consider
useful; the first is no longer future work, and we keep the section to
record what it was and what came of it.

\subsection{A Flat Temporal Index for Unbounded Ranges --- Implemented}
\label{sec:future:global-index}

The GLOBAL-query fix (Section~\ref{sec:discussion:interpretation}) was
already implemented as of the previous revision of this paper; the same
class of gap remained, at that time, for TEMPORAL queries with an
open-ended range (Section~\ref{sec:discussion:temporal-unbounded}): the
year-bucketed temporal index had no bounded candidate set to fall back on,
so it degraded toward consulting every bucket. This section originally
proposed a flat, sorted temporal index (an interval tree, or a sorted
array of (start, end) pairs searched by binary search) alongside the
bucketed one, so that an unbounded query could binary-search to its
starting point and scan forward instead. We have since implemented exactly
that (\texttt{sfdb.sheaf.indexes.TemporalIndex.facts\_ending\_after} and
its symmetric counterpart \texttt{facts\_starting\_before}); the
benchmarked result is a $1.1$--$1.4\times$ SFDB speedup on unbounded
TEMPORAL queries rather than the $\sim 40\times$ slowdown reported
previously, described in full in
Section~\ref{sec:discussion:temporal-unbounded}. We record the original
proposal here rather than deleting it, since the fact that it turned out
to be a small, well-scoped change rather than a genuine architectural
limitation is itself part of the paper's evidence about how easy this
class of gap is to mistake for a design ceiling.

\subsection{Real-World Datasets and Standard Benchmarks}
\label{sec:future:real-data}

The evaluation uses synthetic data. Repeating it on
LUBM~\cite{guo2005lubm} and on a subset of DBpedia or Wikidata would
provide numbers that can be compared against published results for
Jena, GraphDB, and Virtuoso. The LUBM data loader is already implemented
(\texttt{sfdb.datasets.lubm}); the missing piece is wiring it into the
benchmark runner and reporting cross-engine numbers on the LUBM query set.

\subsection{Integration With a Production RDF Store}
\label{sec:future:jena}

The evaluation in Section~\ref{sec:evaluation} compares SFDB only against
our own KG baseline; Section~\ref{sec:discussion:threats} names this as
implementation bias and flags LOOKUP as the query class most likely to
overstate SFDB's advantage against a mature engine. As a small, honest
check on that risk rather than a full answer to it, we ran a single
external-baseline spot-check: the same $1\,000$-fact stream and the same
LOOKUP anchor entity used in Section~\ref{sec:evaluation}, inserted into
\texttt{rdflib}~\cite{rdflib} (a real, independent, in-memory RDF library,
reachable via \texttt{scripts/rdflib\_spotcheck.py}) and queried with the
equivalent SPARQL pattern. SFDB was $12.9\times$ faster than rdflib on
this single class at this single scale, both engines returning the same
$287$ results. We report this as a preliminary, narrowly-scoped data
point, not as validation of the paper's broader claims: it is one query
class, one scale, and rdflib is itself an in-memory, single-process
library rather than a disk-backed, transactionally-serious store like
Jena TDB2, Virtuoso, or GraphDB. Extending the comparison to those systems
at the scales and query classes reported in Section~\ref{sec:evaluation}
--- rather than this one spot-check --- remains the future work needed to
report absolute performance against the state of the art.

\subsection{Incremental Restriction and Global-Section Maintenance}
\label{sec:future:incremental}

Restriction maps are computed on demand and cached. Incremental
maintenance---updating cached restriction maps and the global-section
cache on insert rather than invalidating them---would improve query
latency in workloads that mix inserts and queries.

\subsection{Concurrent and Distributed Execution}
\label{sec:future:distributed}

Global-section computation is embarrassingly parallel across independent
context sub-trees. A parallel implementation would remove the single-node
scalability ceiling. A distributed implementation would additionally
require a partitioning strategy for the context poset and an inter-node
restriction-map protocol.

\subsection{Compression of Contextually Redundant Data}
\label{sec:future:compression}

Section~\ref{sec:eval:memory} reports that SFDB uses $1.3$--$4.8\times$
more resident memory per fact than the KG baseline, because each fact is
registered in roughly eight to ten open sets plus four auxiliary indexes.
Sections that share a common context or that agree under restriction on a
large sub-tree admit context-aware compression schemes that are not
available to a flat triple store: rather than storing a full section
reference per open set, a fact could store one canonical reference plus a
small bitset of the open sets it belongs to, reconstructed on read.
Realising this would directly reduce the memory overhead identified in
Section~\ref{sec:eval:memory}, which is currently the paper's primary
reported cost.

\subsection{Hybrid Storage --- Revisited}
\label{sec:future:hybrid}

An earlier version of this section proposed a hybrid system that would
store facts under both representations and let the optimiser choose the
access path per query, on the premise that this would combine SFDB's
LOOKUP advantage with a GLOBAL advantage that belonged to the KG baseline.
That premise no longer holds: Section~\ref{sec:evaluation} reports SFDB
faster than the KG baseline on GLOBAL as well as LOOKUP, and
Section~\ref{sec:discussion:temporal-unbounded} closes the one query shape
(unbounded TEMPORAL) where the KG baseline had been competitive. With no
latency-motivated reason left to maintain two copies of the data, full
dual-representation storage is not the extension we would prioritise: it
would double the memory cost identified in Section~\ref{sec:eval:memory}
--- the paper's one remaining tradeoff --- to hedge against a latency gap
that the last two fixes in this evaluation have, between them, closed for
every class we measure. A narrower version of the idea could still be
worth revisiting if a workload surfaces a genuine KG-favoured access
pattern this evaluation does not cover, but we would want that pattern
identified first rather than storing everything twice speculatively.
