\section{Implementation}
\label{sec:implementation}

The implementation is a Python 3.12+ package of approximately $15\,000$
lines organised into ten modules and covered by $347$ passing tests. This
section describes the parts of the codebase that a reader would need to see
in order to understand or reproduce the paper's claims.

\subsection{Repository Structure}
\label{sec:impl:repository}

\begin{description}
    \item[\texttt{sfdb.common}] Shared data types (\texttt{SemanticFact},
          \texttt{Entity}, \texttt{Relation}, \texttt{Context}) and the
          engine ABC that both storage engines implement.
    \item[\texttt{sfdb.canonical}] The canonical intermediate
          representation and the bidirectional mapping between it and each
          engine's native format.
    \item[\texttt{sfdb.kg}] The Knowledge Graph engine: dictionary
          encoding, SPO/POS/OSP indexes, reification, and a SPARQL-inspired
          query executor.
    \item[\texttt{sfdb.sheaf}] The sheaf-inspired engine: incremental
          topology, open-set index, stalk index, restriction cache,
          neighbourhood index, global-section constructor, and consistency
          checker.
    \item[\texttt{sfdb.query}] The shared query language, parser, and
          logical plan representation.
    \item[\texttt{sfdb.optimizer}] The shared cost-based optimiser
          (predicate pushdown, dead-operator elimination) with
          engine-specific cost models.
    \item[\texttt{sfdb.datasets}] Synthetic dataset generation and a
          LUBM~\cite{guo2005lubm} loader.
    \item[\texttt{sfdb.benchmark}] The benchmark harness: runners,
          per-engine adapters, metric collectors, and the cross-engine
          verifier.
    \item[\texttt{sfdb.storage}] JSON, MessagePack, and Parquet
          serialisation for facts and results.
    \item[\texttt{sfdb.visualization}] Plot generation used by
          \texttt{scripts/generate\_paper.py} to produce the figures in
          Section~\ref{sec:evaluation}.
\end{description}

\subsection{Mapping to the Formal Model}
\label{sec:mapping:kg-real}
\label{sec:mapping:sh-real}

The KG mapping $\phi_{\mathrm{KG}}$ reifies a fact into a set of triples
against a fresh event identifier. Given a fact of arity $k$, the mapping
emits $k + 4$ triples: a type triple $(e, \texttt{rdf:type}, \texttt{sfdb:Fact})$,
a subject triple, a relation triple, one triple per object slot, a context
triple, and a confidence triple. Provenance and temporal envelopes add
additional triples when present. The inverse mapping $\psi_{\mathrm{KG}}$
gathers all triples sharing the same event identifier and reconstitutes
the original fact.

The sheaf mapping $\phi_{\mathrm{Sh}}$ stores the fact as a single
\texttt{LocalSection} attached to its context. The section retains the
complete tuple of arguments. No decomposition takes place; retrieval by
fact identifier is a single dictionary lookup. The inverse mapping is the
identity on the stored payload.

Both mappings are lossless in the sense of Definition~\ref{def:lossless}.
The round-trip $\psi \circ \phi = \mathrm{id}_\mathcal{F}$ is enforced by
the test suite for every fact inserted during a benchmark run
(\texttt{tests/test\_roundtrip.py}).

\subsection{Configuration and Reproducibility}
\label{sec:impl:config}

The system reads configuration from three sources, in order of increasing
priority: a YAML file (\texttt{configs/benchmark.json} for the default
benchmark suite), environment variables prefixed with \texttt{SFDB\_}, and
programmatic overrides through a \texttt{Config} dataclass. Every
benchmark run seeds Python's PRNG and NumPy's PRNG from the same integer
(default $42$), records the git commit hash, the Python version, the
\texttt{uv.lock} hash, and a checksum of the generated dataset, and stores
them in a per-run reproducibility manifest.

\subsection{Testing}
\label{sec:impl:testing}

The test suite uses \texttt{pytest}. Module tests validate individual
components: topology construction, restriction composition, stalk indexing,
reification round-trips, query planning, and canonicalisation. Cross-engine
tests (\texttt{tests/test\_cross\_engine.py}) execute the same fact stream
and the same query set on both engines and assert canonical equality of the
result sets. Round-trip tests (\texttt{tests/test\_roundtrip.py}) assert
that $\psi \circ \phi = \mathrm{id}$ on both engines for a large synthetic
population. As of the current revision, all $347$ tests pass in $\approx 3$
seconds on a modern laptop.

\subsection{Engine Adapters for External Systems}
\label{sec:impl:adapters}

The benchmark harness supports pluggable engine adapters. Two adapters are
fully implemented and used in the evaluation: \texttt{KGEngineAdapter} and
\texttt{SheafEngineAdapter}. A partial adapter for Apache Jena, backed by
the \texttt{rdflib} Python library, is available but not integrated into
the results reported here because we do not yet claim like-for-like
comparability with a production Jena TDB2 deployment. Scaffolds for
Blazegraph and Neo4j exist but raise \texttt{NotImplementedError} on
construction; they are place-holders for planned integration work.
