\section{Introduction}
\label{sec:introduction}

Knowledge graphs have become the default representation for semantic
information on the web and in enterprise systems. Their foundations are simple:
every fact is a triple $(s, p, o)$ that asserts a binary relationship
between a subject, a predicate, and an object~\cite{manola2004rdf,sparql2013}. The
model is uniform, transparent, and has an enormous accumulated tooling base. It
is also a poor fit for large classes of real data.

Two mismatches recur across domains. First, many facts are not binary. A drug
interaction record carries a drug, a target, a dosage, a route, a study population,
and a confidence interval, all in one statement. A financial transaction pairs
a buyer, a seller, an instrument, a price, a currency, a venue, and a
timestamp. Encoding such facts as triples requires \emph{reification}: an
auxiliary event node is minted, each attribute becomes its own triple pointing
at that node, and reconstructing the original fact costs one join per attribute.
Second, the same fact often holds in some contexts and not in others. A
biological interaction that is valid \emph{in vitro} may fail \emph{in vivo};
a corporate policy that applies in the European subsidiary may not apply in
North America. Context is not first-class in the triple model. It is added
either through separate named graphs~\cite{carroll2005named}, through further
reification, or through ad-hoc conventions in the predicate URI.

Both problems have well-known workarounds, and mature triple stores handle
them competently at scale. The question we pose in this paper is whether there
is anything to gain from a database whose primary organising principle is not
the triple but the \emph{context}. Concretely, we ask whether one can build a
storage engine that (i)~stores each fact as an atomic record, without
reification, and (ii)~indexes those records by the context poset in a way that
answers entity-centric queries via a direct index lookup while remaining verifiably
equivalent to a conventional RDF-style store on the queries that both engines
can answer.

The mathematical vocabulary that fits this question naturally is that of
presheaves on a poset of contexts. A presheaf assigns a set of local
sections---in our setting, facts---to each context, and equips the assignment
with restriction maps that specialise a fact from a broader context to a
narrower one. A sheaf is a presheaf that satisfies two additional axioms of
locality and gluing. We stress at the outset that on a finite poset endowed
with the Alexandrov topology, these axioms hold vacuously. The system we describe
is therefore a \emph{presheaf-based} database in the strict technical sense; the
sheaf terminology is retained because the operational vocabulary of restriction,
stalks, and global sections is what our implementation actually exposes, not
because we claim a non-trivial sheaf-theoretic theorem.

\subsection*{Contributions}
\label{sec:contributions}

The contributions of this paper are, in order:

\begin{enumerate}
    \item A concise formal model of a context-indexed semantic fact store,
          expressed in the language of presheaves on a finite poset
          (Section~\ref{sec:mathematical_model}). The model is deliberately
          minimal: it names only the structure that the implementation uses.

    \item A prototype implementation, the \emph{Semantic Fact Database}
          (SFDB), that realises the model in roughly $15\,000$ lines of
          Python (Section~\ref{sec:implementation}). Two storage engines are
          provided over a shared canonical schema: a reference triple store
          with SPO/POS/OSP indexes and reification for $n$-ary facts, and a
          sheaf-inspired engine with a stalk index, a restriction cache, and
          an on-demand global-section constructor.

    \item A canonical intermediate representation together with a
          bidirectional mapping to and from each engine's native format,
          used to establish semantic equivalence at every insert and every
          query (Section~\ref{sec:correctness}). Cross-engine equivalence is
          enforced by the test suite ($347$ passing tests) and re-checked
          on every benchmark run.

    \item A benchmark evaluation across three query classes (LOOKUP,
          GLOBAL, TEMPORAL) and insert throughput, at three scales
          (Section~\ref{sec:evaluation}), with every query at every scale
          checked for cross-engine result equivalence as part of the same
          run. The sheaf-inspired engine is $20$--$25\times$ faster than the
          KG baseline on LOOKUP queries, $50$--$100\times$ faster on GLOBAL
          scans, $1.2$--$1.6\times$ faster on bounded temporal-interval
          queries, and matches it on inserts ($0.74$--$0.99\times$). The
          tradeoff we report is not latency but memory: SFDB's in-memory
          indexing structures use $1.3$--$4.8\times$ more resident memory per
          fact than the KG baseline's reified-triple storage. We also report
          that tightening the cross-engine verification harness during this
          evaluation caught a real semantic bug in the sheaf engine's LOOKUP
          path (Section~\ref{sec:correctness}), which we take as evidence
          for the verification methodology, not just for the numbers.

    \item A discussion of the design's realistic scope, including the
          honest observation that the sheaf condition is vacuous in our
          setting and therefore not the source of any of the performance
          differences we observe (Section~\ref{sec:discussion}). The
          practical value of the model lies in the context poset, the
          stalk index, and the canonical mapping, not in the sheaf axioms.
\end{enumerate}

\subsection*{What This Paper Is Not}

We do not claim that a sheaf-inspired database replaces production RDF
stores such as Apache Jena, GraphDB, or Virtuoso. Our KG baseline is a
research prototype implemented by the same authors as the SFDB engine, and
we call attention to this limitation throughout: the LOOKUP speedups we
report are best interpreted as measurements of what disciplined
context-indexed storage buys over a strict reification-plus-joins design at
comparable engineering quality, not as claims about the state of the art in
RDF engines. We do not report results on standard benchmarks such as LUBM,
BSBM, or WatDiv; extending the evaluation in that direction is future work.
We do not claim novelty in category theory. The formal machinery we use is
standard.

\subsection*{Paper Organisation}
\label{sec:organization}

Section~\ref{sec:related-work} places the work relative to existing knowledge
graph and category-theoretic database literature. Section~\ref{sec:background}
recalls the finite mathematical background we rely on. Section~\ref{sec:problem}
states the storage and equivalence problems formally.
Section~\ref{sec:mathematical_model} presents the formal model, together
with the definitions and lemmas that back it.
Section~\ref{sec:system-architecture} describes the system architecture, and
Section~\ref{sec:implementation} the implementation. Section~\ref{sec:evaluation}
reports the benchmark results, organised by query class.
Section~\ref{sec:discussion} interprets the results, and
Section~\ref{sec:limitations} states the limitations. Section~\ref{sec:future}
outlines future work, and Section~\ref{sec:conclusion} concludes.
