\section{Knowledge Graph}
\label{sec:knowledge-graph}

The Knowledge Graph engine is a full RDF-style triple store serving as
the baseline for comparison.

\subsection{Triple Storage}
\label{sec:kg:storage}

Facts are stored as subject-predicate-object triples using dictionary
encoding. Three string-to-integer dictionaries map entity URIs, predicate
URIs, and literal values to compact IDs. The core triple table stores
(subject\_id, predicate\_id, object\_id, object\_type).

\subsection{Indexes}
\label{sec:kg:indexes}

Six permutation indexes support fast access:
\begin{itemize}
    \item SPO: subject $\to$ predicate $\to$ object
    \item POS: predicate $\to$ object $\to$ subject
    \item OSP: object $\to$ subject $\to$ predicate
    \item SOP: subject $\to$ object $\to$ predicate
    \item PSO: predicate $\to$ subject $\to$ object
    \item OPS: object $\to$ predicate $\to$ subject
\end{itemize}
Each index is implemented as a nested hash map or sorted structure.

\subsection{Reification}
\label{sec:kg:reification}

For $n$-ary facts, an intermediate event node is created. Given a fact
with arity $k$, the KG generates $k + 3$ triples (event type, subject,
predicate, $k$ objects). Reconstruction requires $k$ joins.

\subsection{Query Engine}
\label{sec:kg:query}

The KG query engine supports SPARQL-inspired operations: triple pattern
matching, joins (nested-loop, hash), filters, projections, sorting,
aggregation, and limits. The query planner translates logical plans into
physical plans using index selection and join ordering.
