\section{Limitations}
\label{sec:limitations}

We describe the design's limitations honestly and in one place so that a
reader can weigh what the paper does and does not claim.

\subsection{Scale and Deployment}
\label{sec:limits:scalability}

The current prototype is single-node and in-memory. All measurements are at
scales of $10\,000$ facts or below; behaviour at larger scales is
extrapolated from the complexity analysis of Section~\ref{sec:complexity}
and not directly measured. Global-section computation in the worst case is
$O(|\mathcal{C}| \cdot N^2)$, which imposes a natural ceiling on the size
of context poset that the system can support on a single node. This
ceiling is compounded by a concretely measured one: Section~\ref{sec:eval:memory}
reports $1.2$--$3.4\times$ higher resident memory per fact than the KG
baseline, so the single-node memory budget is consumed faster by SFDB than
by an equivalent reified store, independent of the gluing worst case.
Distributing the store across nodes would require a partitioning strategy
for the context poset and a distributed restriction-map protocol; both are
outside the scope of this paper.

\subsection{Mathematical Assumptions}
\label{sec:limits:math}

The formal model assumes a finite context poset equipped with the
Alexandrov topology. This is the right setting for the intended workload,
but it is also the setting in which the sheaf condition holds vacuously.
Nothing in the paper claims a non-trivial sheaf theorem. The gluing
operation in the implementation is a defensive consistency check rather
than a step in a proof; in the presence of genuinely contradictory local
data (which our schema currently forbids), the algorithm would report the
conflict rather than resolve it.

\subsection{Implementation Limitations}
\label{sec:limits:implementation}

The prototype does not currently support concurrent transactions, on-disk
persistence beyond the built-in SQLite backing store, distributed
execution, streaming or incremental fact ingestion, or SPARQL compliance.
An earlier draft of this paper described the query language as "a small
subset of SPARQL"; that overstated it. It is a bespoke query DSL whose
surface syntax borrows SPARQL's \texttt{SELECT}/\texttt{WHERE} keywords and
\texttt{?var} notation but does not parse SPARQL's graph-pattern braces or
its \texttt{FROM} clause, together with a few sheaf-native operations
(restriction, stalk lookup). The typed \texttt{Query} object used
internally by both engines and by the benchmark harness
(Section~\ref{sec:evaluation}) does not depend on this text surface. A
real SPARQL front end and a proper transactional layer are engineering
work that would be needed for any real deployment but that would not
change the paper's structural findings.

\subsection{Baseline and Benchmark Limitations}
\label{sec:limits:benchmark}

The KG baseline is a research prototype implemented by the same team as
the sheaf engine. It is a controlled reference point, not a
state-of-the-art RDF store. The baseline is fair for the purpose of
isolating the effect of context-indexed storage because both engines
share the same serialisation, canonicalisation, and storage backend
(SQLite); the only architectural difference is the indexing strategy
(triple indexes vs.\ stalk+context indexes). The evaluation therefore quantifies what
disciplined context-indexed storage buys over a strict reification-plus-joins
implementation at comparable engineering effort. Repeating the evaluation
against a production RDF store on a standard benchmark such as
LUBM~\cite{guo2005lubm}, BSBM, or WatDiv is the natural next step and is
called out as future work. The synthetic datasets, while parameterised,
may not exhibit all pathologies of organic knowledge graph data.
