We present the complete formal mathematical model underlying SheafDB.
This section consolidates notation, definitions, and structural properties.
All definitions are standard within sheaf theory~\cite{maclane1992sheaves}
and are adapted to the finite, discrete setting of knowledge representation.

\subsection{Formal Notation}
\label{sec:notation}

We use the following notation throughout:

\begin{itemize}
    \item $\mathbb{N}$: natural numbers $\{0,1,2,\ldots\}$.
    \item $\mathcal{P}(X)$: power set of $X$.
    \item $|X|$: cardinality of $X$.
    \item $f: A \to B$: total function; $f: A \rightharpoonup B$: partial function.
    \item $g \circ f$: composition $(g \circ f)(x) = g(f(x))$.
    \item $\operatorname{id}_X$: identity function on $X$.
    \item $(P, \leq)$: a poset with order relation $\leq$.
    \item $\downarrow x = \{y \in P \mid y \leq x\}$: down-set of $x$.
    \item $x \wedge y$: meet (greatest lower bound); $x \vee y$: join (least upper bound).
    \item $\mathcal{I}$: set of identifiers (UUIDs).
    \item $\mathcal{E} \subseteq \mathcal{I}$: set of entity identifiers.
    \item $\mathcal{R} \subseteq \mathcal{I}$: set of relation identifiers.
    \item $\mathcal{V} = \mathcal{I} \sqcup \mathbb{L}$: values (identifiers or literals).
    \item $\mathcal{C}$: set of contexts, a finite poset.
    \item $\mathcal{F}$: set of all possible semantic facts.
    \item $F$: a presheaf or sheaf.
    \item $F(U)$: sections over open set $U$.
    \item $\rho_{V,U}: F(V) \to F(U)$: restriction map for $U \subseteq V$.
    \item $F_x$: stalk at point $x$: $\varinjlim_{U \ni x} F(U)$.
    \item $\Gamma(X, F)$: global sections $F(X)$.
    \item $\llbracket Q \rrbracket$: denotation (result) of query $Q$.
\end{itemize}

\subsection{Category-Theoretic Setting}
\label{sec:categorical}

Let $\mathbf{Set}$ denote the category of sets and functions,
and $\mathbf{Pos}$ the category of partially ordered sets and monotone maps.
A presheaf on a small category $\mathcal{C}$ is a functor
$F: \mathcal{C}^{\mathrm{op}} \to \mathbf{Set}$.

In our setting, $\mathcal{C}$ is the context poset $(\mathcal{C}, \leq)$
viewed as a category: objects are contexts, and there is a unique morphism
$c \to d$ iff $d \leq c$. The opposite category $\mathcal{C}^{\mathrm{op}}$
reverses this: a morphism corresponds to $c \leq d$, i.e., restriction from
a broader context to a narrower one.

\input{sections/formal_definitions.tex}
\input{sections/theorems.tex}



