\section{Motivation}
\label{sec:motivation}

We motivate the design by working through a concrete example that recurs, in
various forms, across our target domains. Consider the following statement,
recorded in a biomedical knowledge base:

\begin{quote}
Aspirin, at a dose of $100\text{ mg}$ per day, taken orally by adult subjects,
inhibits COX-1, with a confidence of $0.92$, according to a randomised trial
published in 2019, in the clinical-adult context.
\end{quote}

The statement is a single fact about a single relationship. Its arguments are
the compound (\emph{aspirin}), the dose ($100\text{ mg/day}$), the route
(\emph{oral}), the population (\emph{adult}), the target (\emph{COX-1}), the
outcome (\emph{inhibition}), the confidence ($0.92$), the source (\emph{a
randomised trial published in 2019}), and the context of validity
(\emph{clinical, adult}). There are two things to notice.

\subsection{The Cost of Reifying $n$-ary Facts}
\label{sec:motivation:triples}

Recorded as RDF triples, the statement above cannot fit into a single
$(s, p, o)$ record. The standard workaround is reification: a fresh event node
$e$ is minted, the fact is asserted as $(e, \texttt{rdf:type}, \texttt{Fact})$,
and each argument becomes its own triple pointing at $e$. For a fact with
$k$ arguments, one obtains $O(k)$ triples plus the event-type triple; the
inverse operation, reconstructing the original fact from the triples, requires
$k$ joins on the event node. On our example this yields roughly ten triples in
place of one statement, and any query that touches multiple arguments--- ``all
compounds that inhibit COX-1 at oral doses above $50\text{ mg}$''---must join
across all of them.

Named graphs~\cite{carroll2005named} and property graphs~\cite{angles2017property}
soften this problem in specific ways: named graphs attach a fourth coordinate
(the graph name) to each triple, and property graphs decorate edges with
attribute maps. Both mechanisms improve on pure RDF for particular workloads,
but neither of them removes the underlying atomicity mismatch: the storage
unit is smaller than the semantic unit, and the cost of reassembly is paid
on every query.

\subsection{Contexts Are Not First-Class}
\label{sec:motivation:contexts}

The example also contains a \emph{context}: the statement is asserted \emph{in
the clinical-adult context}. The same clinical fact, restricted to a
paediatric population, may fail. In practice, contexts are usually encoded
either through named graphs (one graph per context), through further
reification (each fact carries a $\texttt{sfdb:context}$ triple), or through
convention in the predicate URI. All three approaches work, but none of them
makes context part of the database's core index structure. Queries that scope
naturally by context--- ``all facts that hold in the \emph{clinical.adult}
context''---end up as filters or joins over auxiliary data rather than as
primary-index lookups.

\subsection{What a Context-Indexed Store Would Buy}
\label{sec:motivation:sheaf}

Suppose that, instead of triples, the primary storage unit were the full
fact---the tuple of all arguments plus the context---and that the primary
index were keyed by context. Two consequences follow immediately. A lookup
that resolves a specific entity within a specific context becomes a single
hash probe into the stalk at that context, returning the complete fact
without joins. A query that scopes to a sub-tree of the context poset touches
exactly the sections attached to that sub-tree, without scanning the whole
store. Neither observation is deep; both are essentially bookkeeping. The
question is whether the bookkeeping pays for itself in practice, and whether
its costs on the queries it does not help---in particular, unrestricted
full-scan queries---are acceptable.

The remainder of the paper answers those questions concretely. Section
\ref{sec:mathematical_model} names the structure precisely and identifies the
minimal presheaf machinery that our implementation actually uses. Sections
\ref{sec:system-architecture} and \ref{sec:implementation} describe the
storage engine that realises it. Section~\ref{sec:evaluation} measures both
the benefit on LOOKUP queries and the cost on GLOBAL queries, on identical
workloads verified for semantic equivalence against a conventional
triple-store baseline.

\subsection{Research Gap}
\label{sec:motivation:gap}

Sheaf theory has been applied to distributed computing, sensor
fusion~\cite{robinson2020sheaf}, and categorical database
theory~\cite{patterson2022sheaf}. Formal Concept Analysis~\cite{ganter1999formal}
uses similar lattice structures for concept hierarchies but does not include
restriction maps or a notion of stalk. To the best of our knowledge, no prior
system combines a context-indexed storage layer with a stalk-based query path
and a mechanised proof of equivalence against a triple-store baseline on the
same core schema.
