\section{Problem Statement}
\label{sec:problem}

\subsection{Informal Description}
\label{sec:problem:informal}

We have a collection of semantic facts. Each fact describes a relationship
between entities within a specific context. Facts are organized by context,
and contexts form a hierarchy from general to specific. The system must
support:

\begin{enumerate}
    \item Storing facts and retrieving them by their properties and context.
    \item Answering queries that may span multiple contexts.
    \item Reconstructing global facts from local observations.
    \item Ensuring semantic equivalence regardless of how facts are stored.
\end{enumerate}

\subsection{Formal Description}
\label{sec:problem:formal}

Let $\mathcal{F}$ be the set of all possible semantic facts (Definition~\ref{def:semanticfact}),
$\mathcal{C}$ a finite poset of contexts, and $Q$ a query language over $\mathcal{F}$.

\textbf{Storage problem:} Given a finite set $F \subseteq \mathcal{F}$,
construct a data structure $\Sigma(F)$ supporting:
\begin{enumerate}
    \item $\textsc{Insert}(f)$: add $f$ to $\Sigma$.
    \item $\textsc{Query}(Q)$: return $\{f \in \Sigma \mid f \text{ matches } Q\}$.
    \item $\textsc{Glue}()$: reconstruct global sections from local ones.
\end{enumerate}

\textbf{Equivalence problem:} For two data structures $\Sigma_1, \Sigma_2$
built from the same $F$, ensure
$\llbracket Q \rrbracket_{\Sigma_1} \cong \llbracket Q \rrbracket_{\Sigma_2}$
for all $Q \in \mathcal{L}(Q)$.

\textbf{Optimization problem:} Minimize query latency $t(Q, \Sigma)$
subject to storage constraints, for expected query distributions.
