\section{Related Work}
\label{sec:related-work}

Our work sits at the intersection of three research traditions. This section
places it relative to each of them and states, plainly, what we do and do not
adopt from the mathematical vocabulary we borrow.

\paragraph{Knowledge graphs and semantic stores.}
The RDF triple model~\cite{manola2004rdf} and its query language
SPARQL~\cite{sparql2013} form the backbone of the modern semantic web. Production
triple stores---Apache Jena, Virtuoso, GraphDB, Blazegraph---have evolved highly
tuned index structures (SPO, POS, OSP, and permutations), custom storage
formats, and cost-based query optimisers for the reified triple representation.
Named graphs~\cite{carroll2005named} extend the triple with a fourth graph
coordinate and are the standard mechanism for scoping. Property
graphs~\cite{angles2017property}, as implemented in Neo4j and Amazon Neptune,
decorate edges with attribute maps, which handles some higher-arity cases
gracefully but reintroduces the reification problem for facts of arity above
two or three.

The reification cost of representing $n$-ary facts in triples is well
documented~\cite{angles2008survey}. It is $O(k)$ triples and $O(k)$ joins per
fact of arity $k$. Various mitigations exist---SPARQL 1.1's property paths,
star-schema optimisations, RDF-star for statement-level annotation---but the
underlying atomicity mismatch remains: the storage unit is smaller than the
semantic unit. Our KG baseline implements the standard reified approach with
SPO/POS/OSP indexes and dictionary encoding. We do not claim this baseline is
state of the art; it is a controlled reference point.

\paragraph{Categorical and topological database models.}
The categorical database
literature~\cite{patterson2022sheaf,spivak2012functorial} recasts schemas as
categories and instances as functors. Sheaf structures have been used for
distributed sensor fusion~\cite{robinson2020sheaf} and for reasoning about
locally consistent but globally inconsistent data collections. Simplicial
complexes and cellular sheaves have been proposed as data models for
higher-order relationships~\cite{edelsbrunner2010computational}. Formal Concept
Analysis~\cite{ganter1999formal} organises data around a lattice of concepts
that shares surface structural features with our context poset, but does not
include the restriction maps or stalks that our design uses operationally.

We borrow terminology and structural intuition from this literature, but
we are not proving a theorem about sheaves. As we make explicit throughout, the
sheaf condition is vacuous on the finite Alexandrov topology that fits our
use case, so anything the model buys us empirically is buying it because of
the presheaf structure and the choice of indexing, not because of the sheaf
axioms.

\paragraph{Context and provenance in semantic data.}
Contexts, named graphs, and provenance annotations have received sustained
attention in the semantic web community~\cite{carroll2005named}. Approaches
range from explicit named graphs, to context-carrying quads, to reified
provenance ontologies such as PROV-O. Our contribution is neither a new
context ontology nor a new provenance model; we adopt a straightforward
dot-separated hierarchy of contexts as our first-class organising principle and
show what a database built around it looks like when the mathematical structure
is taken seriously.

\paragraph{Positioning.}
To the best of our knowledge, no prior system provides all of the following
together: a context-indexed semantic fact storage engine with stalk-based
LOOKUP, a co-implemented reified triple-store baseline, a canonical
intermediate representation with a mechanised proof of semantic equivalence
between the two, and a benchmark evaluation that reports both wins and losses
across a LOOKUP/GLOBAL taxonomy. The individual ingredients are all standard,
in some cases decades old. What this paper contributes is a coherent design
that combines them and an honest evaluation of what the combination is and is
not worth.
