\subsection{Theorems}
\label{sec:theorems}

\subsubsection{Semantic Preservation}
\label{sec:thm:preservation}

\begin{theorem}[Semantic Preservation / Round-trip Correctness]
\label{thm:preservation}
Let $T: \mathcal{F} \to \mathcal{T}(\mathcal{F})$ be the triple decomposition
function (reifying an $n$-ary fact into $2+n$ triples), and
$R: \mathcal{T}(\mathcal{F}) \to \mathcal{F}$ be the reconstruction function.
For any semantic fact $f \in \mathcal{F}$:
\[
R(T(f)) = f.
\]
\end{theorem}

\begin{proof}[Proof Sketch]
Decomposition $T$ creates triples $(s, \mathtt{rdf:type}, \mathtt{rdf:Statement})$,
$(s, \mathtt{rdf:subject}, f.\text{subject})$, $(s, \mathtt{rdf:predicate}, f.\text{relation})$,
and $(s, \mathtt{rdf:object}_j, f.\text{objects}_j)$ for each $j$.
Reconstruction $R$ groups triples by statement ID $s$ and inverts each step.
Since $T$ is injective and $R$ is its left inverse on $\operatorname{im}(T)$,
the round-trip is identity.
\end{proof}

\subsubsection{Functoriality of Restriction}

\begin{theorem}[Restriction Functoriality]
\label{thm:restriction}
The restriction maps $\rho_{d,c}: F(d) \to F(c)$ for $c \leq d$ satisfy:
\[
\rho_{c,c} = \operatorname{id}_{F(c)}, \qquad
\rho_{e,c} = \rho_{d,c} \circ \rho_{e,d} \quad (c \leq d \leq e).
\]
Hence $F: \mathcal{C}^{\mathrm{op}} \to \mathbf{Set}$ is a functor (a presheaf).
\end{theorem}

\subsubsection{Sheaf Axioms}

\begin{theorem}[Locality]
\label{thm:locality}
Let $F$ be a presheaf on $(X, \mathcal{T})$, $U \in \mathcal{T}$,
$\{U_i\}$ an open cover of $U$, and $s, t \in F(U)$.
If $\rho_{U,U_i}(s) = \rho_{U,U_i}(t)$ for all $i$, then $s = t$.
\end{theorem}

\begin{theorem}[Gluing]
\label{thm:gluing}
Let $F$ be a sheaf on $(X, \mathcal{T})$, $U \in \mathcal{T}$,
$\{U_i\}$ an open cover of $U$, and $\{s_i \in F(U_i)\}$ a compatible family:
$\rho_{U_i, U_i \cap U_j}(s_i) = \rho_{U_j, U_i \cap U_j}(s_j)$ for all $i,j$.
Then there exists a unique $s \in F(U)$ with $\rho_{U,U_i}(s) = s_i$ for all $i$.
\end{theorem}

\begin{proof}[Proof Sketch]
For each $x \in U$, pick $i$ with $x \in U_i$ and define $s(x) = s_i(x)$.
Compatibility ensures this is well-defined on overlaps.
Uniqueness follows from the locality axiom.
\end{proof}

\subsubsection{Canonical Equivalence}

\begin{theorem}[Invertibility of the Canonical Mapping]
\label{thm:canonical}
Let $\phi: \mathcal{F} \to \mathfrak{F}$ be the canonical mapping from facts to canonical facts,
and $\psi: \mathfrak{F} \to \mathcal{F}$ the inverse mapping. Then:
\[
\psi(\phi(f)) = f,\quad \phi(\psi(\mathfrak{f})) = \mathfrak{f}
\]
for all $f \in \mathcal{F}$, $\mathfrak{f} \in \mathfrak{F}$.
\end{theorem}

\begin{theorem}[Cross-Engine Query Equivalence]
\label{thm:equivalence}
Let $\Sigma_{\mathrm{KG}}$ and $\Sigma_{\mathrm{Sheaf}}$ be knowledge states
derived from the same canonical model $\mathfrak{F}$.
Let $Q$ be a canonical query. Then:
\[
\llbracket Q \rrbracket_{\Sigma_{\mathrm{KG}}} \cong \llbracket Q \rrbracket_{\Sigma_{\mathrm{Sheaf}}}
\]
where $\cong$ denotes equality up to identifier isomorphism.
\end{theorem}

\subsubsection{Structural Properties}

\begin{theorem}[Context Meet as Open Set Intersection]
\label{thm:meet}
For any $c_1, c_2 \in \mathcal{C}$ in the Alexandrov topology $\mathcal{T}_A$:
\[
U_{c_1} \cap U_{c_2} = U_{c_1 \wedge c_2}.
\]
\end{theorem}

\begin{proof}
$d \in U_{c_1} \cap U_{c_2}$ iff $d \leq c_1$ and $d \leq c_2$, which is
equivalent to $d \leq c_1 \wedge c_2$, i.e., $d \in U_{c_1 \wedge c_2}$.
\end{proof}

\begin{theorem}[Preservation of Referential Integrity]
\label{thm:integrity}
For any knowledge state $\Sigma$ and fact $f \in \Sigma$:
\[
\operatorname{subject}(f) \in \mathcal{E}_\Sigma,\quad
\operatorname{relation}(f) \in \mathcal{R}_\Sigma,\quad
\forall \omega_j \in \vec{o}: \omega_j \in \mathcal{E}_\Sigma \text{ if reference}.
\]
\end{theorem}

\begin{theorem}[Determinism of Query Execution]
\label{thm:determinism}
For a fixed query $Q$ and fixed knowledge state $\Sigma$,
the result $\llbracket Q \rrbracket_\Sigma$ is uniquely determined,
independent of execution order, caching state, parallelism, or query history.
\end{theorem}

\begin{theorem}[Consistency of Global Sections]
\label{thm:global}
A family of local sections $\{s_i \in F(U_i)\}$ can be glued to a global section
$s \in F(\mathcal{C})$ iff for every pair $(i,j)$:
\[
\rho_{U_i, U_i \cap U_j}(s_i) = \rho_{U_j, U_i \cap U_j}(s_j).
\]
\end{theorem}
